Embedding Intuitionistic into Classical Logic

### 2022-10-06 FoSSaCS/fossacs-EIICL.tex
Incorporated previous work

\documentclass[runningheads]{llncs}
\usepackage{amsmath, amsfonts}
\usepackage{wrapfig}
\usepackage{tikz-qtree}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}
% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%

%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\begin{document}
	%
	\title{Embedding Intuitionistic into Classical Logic}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	\author{Alexander Pluska\inst{1}\orcidID{0000-0002-7709-3335} \and
		Florian Zuleger\inst{2,3}\orcidID{0000-0003-1468-8398}}
	%
	\authorrunning{A. Pluska and F. Zuleger}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{TU Wien, Vienna, Austria}
	%
	\maketitle
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic. Curiously the reverse direction has not been covered in literature. We present an effective embedding from intuitionistic into classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
Furthermore, in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas. As part of our construction, we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.

@@

We hope that our embeddings of classical into intuitionistic logic will allow leveraging the tremendous progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades. Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
\end{abstract}

@@

Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g. by assuming non-existence and deriving a contradiction.
The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry-Howard correspondence~\cite{howard1980formulae}.

The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g. the
ILTP library website~\cite{iltp}).
However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g. TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate. Furthermore all the most popular first-order utilize classical logic.
This difference can be partially explained by fundamental differences between the logics.
First of all, determining intuitionistic validity is computationally harder, i.e. in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e. to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} use the naive inverse method, i.e. directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization. This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e. SAT-solvers.

With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
To this end, we give an embedding from intuitionistic logic into classical logic. That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case. In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid. Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent. This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e. if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$. In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
%
In addition to the translation of formulas we give an effective translation of counter-models, i.e. for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.

In summary, our work suggests a new approach for obtaining a constructive proof using a classical prover, that is
\begin{itemize}
	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$,
	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model,
	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
\end{itemize}

\textbf{Related Work.}
Several approaches to obtain constructive content from classical proofs have been studied in the literature.
An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other settings have been studied~\cite{schwichtenberg}.
Another important direction is to equip classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.

\begin{theorem}
\label{thm:reduction-propositional}
	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$. Obtain $\mathcal S^\#$ by including the following formulas:
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
	\end{itemize}
Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence. The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$. Futhermore, there is an effective procedure for translating counter-models between the sentences.
\end{theorem}

@@

In this section we fix notation and recall the semantics for classical and intuitionistic logic.

@@

\subsection{Skolemization and Herbrandization}

@@

\section{IPC to CQC}

@@

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
	\begin{itemize}
		\item $A^{u} = A(u)$ for every propositional variable $A$.
		\item $\bot^u = \bot$.
		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
	with e.g.
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
	where $u$ is some free variable.
\end{definition}

\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:

\begin{lemma}
	$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
\end{lemma}

\section{Normal form translation}
\label{section:tseytin}

Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying. As a first step towards obtaining a propositional formula we want to apply Herbrandization. Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone. A similar translation for intuitionistic propositional logic was also proposed in~\cite{statman1979intuitionistic}.

\begin{definition}
Let $\varphi$ be some formula. For each subformula $\psi$ of $\varphi$, consider some new $n$-ary relation symbol $P_\psi$, where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$. We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order. We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
	\begin{itemize}
		\item For every atomic formula $A$:\\
$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$:\\
$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
		\item For every implication $\varphi \to\psi$:\\
$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$:\\
$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
	\end{itemize}
	%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
\end{definition}

\begin{lemma}
	$\bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)\to\varphi$ and $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi(\vec z_\varphi)$ are valid both classically and intuitionistically.
\end{lemma}

It is straight-forward to extend structures to the additional symbols such that the formulas $S^+(\varphi)\cup\mathcal S^-(\varphi)$ are satisfied, i.e. for each structure $\mathcal M$ and formula $\varphi$ we can define a structure $\mathcal S(\mathcal M, \varphi)$, whose exact definition can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
\begin{lemma}\label{thm:struct1}
For every formula $\varphi$ and structure $\mathcal M$, we have that $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
\end{lemma}

\begin{lemma}\label{thm:struct2}
	For every structure $\mathcal M$,
	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
\end{lemma}

Then from the previous three Lemmata we get:

\begin{corollary}\label{equivalid}
	In both intuitionistic and classical logic
	\begin{itemize}
		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is, and
		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
	\end{itemize}
\end{corollary}

\section{IPC to simplified CQC}

We are now ready to combine the previously presented reductions.
Let $\varphi$ be some propositional formula. As a first step we perform the normalization from the last section, i.e. instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi.$$

>>>>

\documentclass[runningheads]{llncs}
\usepackage{amsmath, amsfonts}
\usepackage{wrapfig}
\usepackage{tikz-qtree}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\begin{document}
	%
	\title{Embedding Intuitionistic into Classical Logic}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	\author{Alexander Pluska \and Florian Zuleger}
	%
	\authorrunning{A. Pluska and F. Zuleger}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{TU Wien, Vienna, Austria}
	%
	\maketitle
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic. Curiously the reverse direction has not been covered in literature. Based on the work in~\cite{claessen2015sat} we establish a small model property for intuitionistic propositional logic which allows us to utilize its Kripke semantics to give a simple embedding into classical propositional logic and quantified boolean formulas. We then transfer this procedure to first-order logic which allows us to give a embedding of intuitionistic first-order logic into classical first-order-logic.

@@

This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity. A first benchmark shows that our approach is competitive. Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
\end{abstract}

@@

Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g. by assuming non-existence and deriving a contradiction.
The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry--Howard correspondence~\cite{howard1980formulae}.

The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g. the ILTP library website~\cite{iltp}).
However, intuitionistic provers perform poorly compared to their classical counterparts, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g. TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate. Furthermore all the most popular first-order utilize classical logic.
This difference can be partially explained by fundamental differences between the logics.
First of all, determining intuitionistic validity is computationally harder, i.e. in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e. to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic.
The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, i.e. direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search. More recently connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic, and show promise.
There have also been some successful attempts to study intuitionistic validity via embedding into higher-order classical logic~\cite{LEO}.
Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic.

With this work we want to examine a new but evident approach for directly leveraging the progress in classical reasoning for intuitionistic logic.
To this end, we give an embedding from intuitionistic logic into classical logic. That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case. In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid. Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent. This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e. if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$. In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
%
In addition to the translation of formulas we give an effective translation of counter-models, i.e. for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.

In summary, our work suggests the following approach:
\begin{itemize}
	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$,
	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model,
	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
\end{itemize}

@@

In this section we fix notation and recall the semantics for classical and intuitionistic logic. Futhermore we recapitulate some results from~\cite{otten2005clausal}.

@@

\subsection{Skolemization and Herbrandization}\label{section:herbrandiaztion}

@@

\subsection{Intuitionistic normal forms}

Finally we note the results from~\cite{otten2005clausal} which are of importance to us.

\begin{lemma}\label{lemma:propositional-normal-form}
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R$ of the form
	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$(A\to B)\to C$$
	called \emph{implication} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$. In general they will contain additional propositional variables.
\end{lemma}

A similar result holds in the first order case.

\begin{lemma}\label{lemma:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function and relation symbols.
\end{lemma}

Since we could not find a proof of this result one is given in the appendix\ref{proof:first-order-normal-form}.

\begin{theorem}
\label{thm:reduction-propositional}
	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$. Obtain $\mathcal S^\#$ by including the following formulas:
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
	\end{itemize}
Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence. The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$. Futhermore, there is an effective procedure for translating counter-models between the sentences.
\end{theorem}

\section{Bounded Intuitionistic Counter-models}

@@

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
	\begin{itemize}
		\item $\bot^u = \bot$.
		\item $A^{u} = A(u)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
	with e.g.
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{r}$$
	where $r$ is a new constant symbol.
\end{definition}

\noindent As all we have done is modeling Kripke frames as a first-order theory we directly obtain:

\begin{lemma}
	$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
\end{lemma}

This is somewhat unsatisfying as we end up in a much more complex logic. However using the normal form from Lemma~\ref{lemma:propositional-normal-form}, i.e. assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
\begin{itemize}
	\item For flat clauses $((A_1\wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_m))^u$ is
	$$\forall w(u\preceq w\to(A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u))).$$
	\item For implication clauses $((A\to B)\to C)^u$ is
	$$\forall w(u\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
\end{itemize}

and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to r\preceq s\to\left(\bigwedge\mathcal R^{(u)S}\wedge \bigwedge\mathcal X^{(u)S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{(u)S}$ is of the form
$$\forall w(s\preceq w\to (A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u)))$$
and each formula $\psi\in\mathcal X^{(u)S}$ is of the form
$$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
with $f_\psi$ being a new function symbol.

We now argue that if a counter-model exists for $\varphi^\circ$ then a small counter-model exists. This is ultimately what enables our translations.

\begin{definition}
For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{(u)S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
		m,&\text{ if $\psi$ is fulfilled at $u$,}\\
		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
	\end{cases}$$. Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
		\item Interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $[\psi_1]\dots[\psi_n]$,}\\
			\psi_1\dots\psi_n\psi, &\text{else.}			
		\end{cases}$$
		\item For propositional variables $P$, we set $$P^{I_T}\left(\psi_1\dots \psi_n\right) = P^I\left(g_{\psi_n}\left(\dots\left(g_{\psi_1}\left(s^I\right)\right)\dots\right)\right).$$
	\end{itemize}
\end{definition}
\begin{lemma}~\label{thm:prop-countermodel-reduction}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq m$.
	\end{enumerate}
\end{lemma}

A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}. Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_\psi$, we may add the assumption that $f_\psi(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

\begin{corollary}\label{cor:prop-tree-model}
	If $\mathcal M$ is a counter-model to $\varphi^{CH}$ then so is $\mathcal M_T$.
\end{corollary}

\begin{remark}
	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic. Our contribution is giving an explicit size and geometry that depends only on the structure of the considered formula.
\end{remark}

\begin{corollary}
	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} in intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
\end{corollary}
